### 2021-04-12 tex/paper.tex
Added TeX source of paper and updated .gitignore to ignore files generated during LaTex build

\documentclass[12pt]{amsart}
\frenchspacing \mathsurround=1pt \emergencystretch=5pt
\tolerance=400

\topmargin = -0.1in \evensidemargin = 0.2in \oddsidemargin = 0.2in
\textheight = 23cm \headheight = 10pt
\textwidth = 15.7cm

\usepackage{amsmath, amsthm, latexsym, amssymb, amsfonts, epsfig, epsf, xcolor, hyperref}
\usepackage{mathtools}
\usepackage{bm,xcolor,tikz,hyperref}
%\usepackage{refcheck}
\usepackage{bm}
\usepackage{blkarray}
\usepackage{booktabs}
\usepackage{enumitem}

\newenvironment{pf}{\proof[\proofname]}{\endproof}
\theoremstyle{plain}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{example}[theorem]{Example}
\newtheorem{remark}[theorem]{Remark}

\newcommand{\rmv}[1]{}

\begin{document}
\title{Coding theory package for Macaulay2}
%%%%%%%%%%%%%%%%%%%%%%%
\author{Taylor Ball}
\address[Taylor Ball]{Department of Mathematics\\University of Notre Dame\\ Notre Dame, IN USA}
\email{trball13@gmail.com}
%%%%%%%%%%%%%%%%%%%%%%%
\author{Eduardo Camps}
\address[Eduardo Camps]{Escuela Superior de F\'isica y Matem\'aticas \\ Instituto Polit\'ecnico Nacional\\ Mexico City, Mexico}
\email{camps@esfm.ipn.mx}
%%%%%%%%%%%%%%%%%%%%%%%
\author{Henry Chimal-Dzul}
\address[Henry Chimal-Dzul]{Department of Mathematics and Center of Ring Theory and its Applications\\ Ohio University\\ Athens, OH USA}
\email{hc118813@ohio.edu}
%%%%%%%%%%%%%%%%%%%%%%%
\author{Delio Jaramillo-Velez}
\address[Delio Jaramillo-Velez]{
Departamento de
Matem\'aticas\\
Centro de Investigaci\'on y de Estudios
Avanzados del
IPN\\
Apartado Postal
14--740 \\
07000 Mexico City, D.F.
}

\email{djaramillo@math.cinvestav.mx}
%%%%%%%%%%%%%%%%%%%%%%%
\author{Hiram H. L\'opez}
\address[Hiram H. L\'opez]{Department of Mathematics and Statistics\\ Cleveland State University\\ Cleveland, OH USA}
\email{h.lopezvaldez@csuohio.edu}
\thanks{}
%%%%%%%%%%%%%%%%%%%%%%%
\author{Nathan Nichols}
\address[Nathan Nichols]{School of Mathematics \\ University of Minnesota Twin Cities \\ Minneapolis, MN USA}
\email{nathannichols454@gmail.com}
%%%%%%%%%%%%%%%%%%%%%%%
\author{Matthew Perkins}
\address[Matthew Perkins]{Department of Mathematics\\ Cleveland State University\\ Cleveland, OH USA}
\email{m.r.perkins73@vikes.csuohio.edu}
%%%%%%%%%%%%%%%%%%%%%%%
\author{Ivan Soprunov}
\address[Ivan Soprunov]{Department of Mathematics\\ Cleveland State University\\ Cleveland, OH USA}
\email{i.soprunov@csuohio.edu}
%%%%%%%%%%%%%%%%%%%%%%%
\author{German Vera-Mart\'inez}
\address[German Vera-Mart\'inez]{Escuela Superior de F\'isica y Matem\'aticas \\ Instituto Polit\'ecnico Nacional\\ Mexico City, Mexico}
\email{gveram1100@alumno.ipn.mx}
%%%%%%%%%%%%%%%%%%%%%%%
\author{Gwyn Whieldon}
\address[Gwyn Whieldon]{}
\email{gwyn.whieldon@gmail.com}
%%%%%%%%%%%%%%%%%%%%%%%

\keywords{Linear codes, locally recoverable codes, Cartesian codes, evaluation codes, Hamming codes}
\subjclass[2010]{Primary 94B05; Secondary 13P25, 14G50, 11T71}

\begin{abstract}
In this Macaulay2 \cite{M2} package we define an object called {\it linear code}. We implement functions that compute basic parameters and objects associated with a linear code, such as generator and parity check matrices, the dual code, length, dimension, and minimum distance, among others. We define an object {\it evaluation code}, a construction which allows to study linear codes using tools of
algebraic geometry and commutative algebra. We implement functions to generate important families of linear codes such as Hamming codes, cyclic codes, Reed--Solomon codes, Reed--Muller codes, Cartesian codes, monomial--Cartesian codes, and toric codes. In addition, we define functions for the syndrome decoding algorithm and locally recoverable code construction, which are important tools in applications of linear codes. The package \textit{CodingTheory.m2} is available at \url{https://github.com/Macaulay2/Workshop-2020-Cleveland/tree/CodingTheory/CodingTheory}
\end{abstract}

\maketitle
\markleft{Ball, Camps, Chimal-Dzul, Jaramillo-Velez, L\'opez, Nichols, Perkins, Soprunov, Vera, Whieldon}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
Coding theory has been extensively studied since 1949, when Claude Shannon proved in his seminal paper \cite{Shannon} that linear codes can be used to reliably transmit information from a single source to a single receiver through a noisy channel. Since then, coding theory has found many important engineering applications. For example, coding theory has been used in designing reliable data storage systems, radio communication protocols, and in the emerging field of quantum computers. Coding theory has close ties with many areas in mathematics including linear algebra, commutative algebra, algebraic geometry, and combinatorics.

In this note we introduce a new package written for Macaulay2 \cite{M2} called \textit{CodingTheory}. The goal of this package is to provide a range of functions for constructing  linear and evaluation codes over finite fields, and for computing some of their main properties. To this aim, we define two objects, namely \textit{linear code} and \textit{evaluation code}. The package also includes implementation of functions for generating important families of linear codes like Hamming codes, cyclic codes, Reed-Solomon codes, Reed-Muller codes, Cartesian codes, monomial-Cartesian codes and toric codes. It also has functions for the syndrome decoding algorithm and locally recoverable codes.

The organization of this note is as follows. In Section~\ref{def} we describe different ways to define a linear code over a finite field using the \textit{CodingTheory} package. In Section~\ref{pro} we show how to compute the main parameters of a linear code: length, dimension, and minimum distance. We also illustrate how to compute some of the main algebraic objects associated with linear codes like generator and parity check matrices, dual codes, etc. In Section~\ref{eva} we give a brief introduction to evaluation codes and describe some functions implemented to study these objects. In Section~\ref{families} we explain how to create some of the most studied families of linear codes, including Hamming codes, cyclic codes, Reed-Solomon codes, and Reed-Muller codes. Finally, we give instructions on how to create locally recoverable codes.

In this paper we do not attempt to fully explain every function distributed in this package.
%It is important to mention that the functions are not fully explained in this paper. 
For a detailed explanation of all functions in the package, we refer to the Macaulay2 help page which can be accessed by
running 
%installPackage("CodingTheory", RemakeAllDocumentation=>true)
\begin{verbatim}
viewHelp CodingTheory
\end{verbatim}

More information about basics of coding theory can be found in \cite{huf-pless,MacWilliams-Sloane,van-lint}. Constructions of codes using commutative algebra as evaluation codes can be seen in
\cite{carvalho4,GLS,Ha1,LSc,MPV,MPV2,algcodes,renteria-tapia-ca2, Ru, SoSo, So1}. Excellent references for theory of vanishing ideals and their properties are \cite{CLO,monalg}.

\section{Defining linear codes}\label{def}
Let $\mathbb{F}_q$ be a finite field with $q$ elements. Mathematically, a {\it linear code} is defined as a vector subspace $C\subseteq \mathbb{F}_q^n.$ For Macaulay2 (M2), a linear code is a submodule of $\mathbb{F}_q^n.$ Assume $q=p^r,$ where $p$ is a prime number and $r$ a positive integer. By definition, the dual code $C^\perp$ is the orthogonal complement of $C$ in $\mathbb{F}_q^n$ with respect to the standard inner product. One can define $C$ by specifying a list $L$ of elements of $\mathbb{F}_q^n$ that span $C$ or by giving a
a {\it generator matrix} $G$ whose rows form a basis of $C$. Alternatively, one can specify a list $L_{H}$ of elements of $\mathbb{F}_q^n$ that span the dual  code $C^\perp$ or a {\it parity check matrix} $H$ whose columns form a basis of the dual code $C^\perp$.
Below are the commands for the constructor linearCode to construct these equivalent instances of the LinearCode type:
\begin{itemize}
\item linearCode($\mathbb{F}_q$,$L$)
%\item linearCode($\mathbb{F}_q^n$, $L$)
%\item linearCode($\mathbb{F}_q^n$, $L_H$, ParityCheck $=>$ true)
%\item linearCode($\mathbb{F}_q$,$L_{H}$, ParityCheck $=>$ true)
\item linearCode($\mathbb{F}_q$,$n$,$L$)
\item linearCode($G$)
\item linearCode($\mathbb{F}_q$,$L_H$, ParityCheck $=>$ true)
\item linearCode($H$, ParityCheck $=>$ true)
\item linearCode($p$,$r$,$n$,$L$)
\item linearCode($p$,$r$,$n$,$L_H$, ParityCheck $=>$ true)
\end{itemize}

Now, here is a more specific example of how to construct a simple linear code:

\begin{example}
$\,$
\begin{verbatim}
i2 : F = GF 4;
i3 : L = {{1,1,0,0},{0,0,1,1}};
i4 : C = linearCode(F,L)
o4 = Code with Generator Matrix: | 1 1 0 0 |
                                 | 0 0 1 1 |
o4 : LinearCode
\end{verbatim}
\end{example}

One way to refer to a primitive element of a finite field is by specifying a symbol using the Variable option of the constructor GF.

\begin{example}
$\,$
\begin{verbatim}
i2 : F = GF(9,Variable => a);
i3 : LH = {{1,0,a,0,0},{0,a,a+1,1,0},{1,1,1,a,0}};
i4 : C = linearCode(F,LH,ParityCheck => true)
o4 = Code with Generator Matrix: | a-1 0 a+1 1 0 |
                                 | 0   0 0   0 1 |
o4 : LinearCode
\end{verbatim}
\end{example}
To construct a linear code from a matrix, it is necessary to correctly specify the underlying field. This can be done by passing a field to the matrix constructor.
\begin{example}
$\,$
\begin{verbatim}
i2 : F = GF 4;

i3 : M = matrix(F, {{1,0,1,0},{0,1,1,1}});

             2       4
o3 : Matrix F  <--- F

i4 : C = linearCode(M)

o4 = Code with Generator Matrix: | 1 0 1 0 |
                                 | 0 1 1 1 |

o4 : LinearCode

\end{verbatim}
\end{example}

\section{Basic parameters linear codes}~\label{pro}

The {\it dimension\/} and the {\it length\/} are two of the basic parameters of a code $C\subseteq \mathbb{F}_q^n$. They are defined as the subspace dimension $\dim_{\mathbb{F}_q} C$ and the ambient space dimension $n$, respectively. A third basic parameter is the {\it minimum weight}, which is given by 
\[
\min\{\|c\|
\colon  c \in C, c \neq 0\},
\]
where $\| c \|$ is the number of non-zero entries of $c$. The {\it rate} of $C$ is defined as the rational number $k/n.$ %The dual of $C$ is defined as the orthogonal complement of $C$ in $\mathbb{F}^n_q$ with respect to the usual dot product in $\mathbb{F}^n_q$.
 Some of the functions that can be used in M2 to compute basic parameters and algebraic objects associated with linear codes are the following: \newline
\begin{minipage}[t]{0.33\textwidth}
\begin{itemize}[leftmargin=1 cm]
\item C.GeneratorMatrix
\item C.Generators
\item C.ParityCheckMatrix
\item C.AmbientModule
\item alphabet C
\end{itemize}
\end{minipage}
\begin{minipage}[t]{0.33\textwidth}
\begin{itemize}
\item field C
\item informationRate C
\item ambientSpace C
\item length C
\item dim C
\end{itemize}
\end{minipage}
\begin{minipage}[t]{0.33\textwidth}
\begin{itemize}
\item minimumWeight C
\item codewords C
\item dualCode C
\item shorten(C, List)
\item ==
\end{itemize}
\end{minipage}

\begin{example}
$\,$
\begin{verbatim}
i2: F = GF 4;
i3 : L = {{1,1,0,0},{0,0,1,1}};
i4 : C = linearCode(F,L)
o4 = Code with Generator Matrix: | 1 1 0 0 |
                                 | 0 0 1 1 |
o4 : LinearCode
i5 : length C
o5 = 4
i6 : dim C
o6 = 2
i7 : informationRate C
     1
o7 = -
     2
o7 : QQ
i8 : ambientSpace C
      4
o8 = F
o8 : F-module, free
i9 : alphabet C
o9 = {0, a, a + 1, 1}
o9 : List
i10 : minimumWeight C
o10 = 2
i11 : dualCode C
o11 = Code with Generator Matrix: | 1 1 0 0 |
                                  | 0 0 1 1 |

o11 : LinearCode
\end{verbatim}
\end{example}

\section{Evaluation codes}\label{eva}
Let $\mathcal{X}=\left\{{\bf a}_1,\ldots,{\bf a}_n\right\}$ be a subset of an $m$-dimensional space $\mathbb{F}_q^{m}$.
Consider a finite dimensional subspace $S\subset {\mathbb{F}_q}[X_1,\ldots,X_m]$  of the $m$-variate polynomial ring over $\mathbb{F}_q$.
%and $S_{\leq d}$ the ${\mathbb{F}_q}$-vector space of all polynomials in $S$ of degree at most $d.$

The {\it evaluation map\/}
\begin{equation*}\label{functionevaluation}
{\rm ev}_S\colon S\longrightarrow {\mathbb{F}_q}^{|\mathcal{X}|},\ \ \ \ \ 
f\mapsto \left(f({\bf a}_1),\ldots,f({\bf a}_n)\right),
\end{equation*}
defines a linear map of ${\mathbb{F}_q}$-vector spaces. The image of ${\rm ev}_S$ in ${\mathbb{F}_q}^{|\mathcal{X}|}$, denoted by $C_{\mathcal{X}}(S)$, %defines a ${\mathbb{F}_q}$-vector subspace. We refer to $C_{\mathcal{X}}(d)$ 
is the {\it evaluation code\/} on the set $\mathcal{X}$ corresponding to $S$. The {\it vanishing ideal} of $\mathcal{X}$, denoted by $I(\mathcal{X})$, is the ideal in $S$ of all polynomials that vanish on  $\mathcal{X}.$ A key observation that allows the use of commutative algebra in studying evaluation codes is that the kernel of the evaluation map ${\rm ev}_S$ is precisely $S\cap I(\mathcal{X})$.

An evaluation code $C_{\mathcal{X}}(S)$ is defined in M2 as a separate type because there are more objects associated with it than with a linear code. For instance, the vanishing ideal associated to the set $\mathcal{X}$ plays an important role when finding and estimating parameters of the code, so it is convenient to be able to access it. Given an evaluation code $C$ in M2, the object C.linearCode is a linear code in M2. The command $\text{evaluationCode}(\mathbb{F}_q,\mathcal{X},L)$ defines an evaluation code where $\mathcal{X}$ is a list of elements in $\mathbb{F}_q^{m}$ and $L$ is a list of polynomials that span $S$. In the case when polynomials in $L$ are monomials, one may give the matrix of exponent vectors instead of $L$.

There are many construction of evaluation codes for specific choices of the set $\mathcal{X}$ and the subspace $S$. These include
Reed-Muller codes, Cartesian and monomial Cartesian codes, toric codes, and evaluation codes from graphs. We refer to 
\cite{carvalho4, Ha1, LSc, lopez-villa, MPV, algcodes, Ru, SoSo} for details on how these codes are defined and what properties they have from coding theory, commutative algebra, and algebraic geometry perspectives. Some functions defined in this package for various constructions of evaluation codes and associated algebraic objects are the following:

\medskip

\begin{minipage}[t]{0.52\textwidth}
\begin{itemize}[leftmargin=0.55 cm]
\item evaluationCode($\mathbb{F}_q$,List,List)
\item toricCode($\mathbb{F}_q$,Integer matrix)
\item cartesianCode($\mathbb{F}_q$,List,List)
\item orderCode($\mathbb{F}_q$,List,List,ZZ)
\item evCodeGraph($\mathbb{F}_q$,Incident Matrix,integer)
\end{itemize}
\end{minipage}
\begin{minipage}[t]{0.45\textwidth}
\begin{itemize}[leftmargin=1 cm]
\item vNumber(Ideal)
\item footPrint(Integer,Integer,Ideal)
\item hYpFunction(Integer,Integer,Ideal)
\item gMdFunction(Integer,Integer,Ideal)
\item vasFunction(Integer,Integer,Ideal)
\end{itemize}
\end{minipage}

\medskip

The mathematical definitions of the vNumber, the footprint function, the hyp function, the generalized footprint function and the Vasconcelos function can be found in~\cite{CSTVV}. The following example shows how to construct an evaluation code.

\begin{example}
$\,$
\begin{verbatim}
i2 : F=GF(4); R=F[x,y,z];
i4 : P={{0,0,0},{1,0,0},{0,1,0},{0,0,1},{1,1,1},{a,a,a}};
i5 : S={x+y+z,a+y*z^2,z^2,x+y+z+z^2};
i6 : C=evaluationCode(F,P,S)
o6 = Code with Generator Matrix: | 0 1 1 1 1   a   |
                                 | a a a a a+1 a+1 |
                                 | 0 0 0 1 1   a+1 |
                                 | 0 1 1 0 0   1   |
o6 : EvaluationCode
i7 : length C.LinearCode
o7 = 6
i8 : dim C.LinearCode
o8 = 3
i9 : C.Points
o9 = {{0, 0, 0}, {1, 0, 0}, {0, 1, 0}, {0, 0, 1}, {1, 1, 1}, {a, a, a}}
o9 : List
i10 : C.VanishingIdeal;
o10 = Ideal of R
\end{verbatim}
\end{example}
\section{Families of linear codes}\label{families}
We continue with the same notation: $n$ represents the length of the code, $k$ the dimension and $q$ the size of the field.
Some families of linear codes that have been implemented in this package are the following:

\medskip

\begin{minipage}[t]{0.5\textwidth}
\begin{itemize}[leftmargin=1 cm]
\item HammingCode($q$,integer)
\item randLDPC($n$, $k$, integer, integer)
\item cyclicCode($\mathbb{F}_q$ ,polynomial, $n$)
\item quasiCyclicCode($\mathbb{F}_q$,list)
\item RSCode($\mathbb{F}_q$,List,integer)
\item RMCode($q$,integer,integer)
\end{itemize}
\end{minipage}
\begin{minipage}[t]{0.5\textwidth}
\begin{itemize}[leftmargin=1 cm]
\item zeroCode($\mathbb{F}_q$,$n$)
\item universeCode($\mathbb{F}_q$,$n$)
\item repetitionCode($\mathbb{F}_q$,$n$)
\item zeroSumCode($\mathbb{F}_q$,$n$)
\item random($\mathbb{F}_q$, $n$, $k$)

\end{itemize}
\end{minipage}

\medskip

Mathematical definitions of the above families can be found in \cite{huf-pless,MacWilliams-Sloane,van-lint}.

\begin{example}
$\,$
\begin{verbatim}
i2 : C = HammingCode(2,3)
o2 = Code with Generator Matrix: | 1 1 1 1 0 0 0 |
                                 | 0 1 1 0 1 0 0 |
                                 | 1 0 1 0 0 1 0 |
                                 | 1 1 0 0 0 0 1 |
o2 : LinearCode
i3 : F=GF(5); R=F[x]; g=x-1; C=cyclicCode(F,g,6)
Cyclic Code
o6 = Code with Generator Matrix: |-1  1  0  0  0  0 |
                                 | 0 -1  1  0  0  0 |
                                 | 0  0 -1  1  0  0 |
                                 | 0  0  0 -1  1  0 |
                                 | 0  0  0  0 -1  1 |
o6 : LinearCode
i7 : F = GF(5);
i8 : L = apply(toList(1..2),j-> apply(toList(1..4),i-> random(F)))
o8 = {{0, -2, 2, -1}, {-2, 1, 0, -1}}
o8 : List
i9 : C=quasiCyclicCode(L)
o9 = Code with Generator Matrix: |  0 -2  2 -1 |
                                 | -1  0 -2  2 |
                                 |  2 -1  0 -2 |
                                 | -2  2 -1  0 |
                                 | -2  1  0 -1 |
                                 | -1 -2  1  0 |
                                 |  0 -1 -2  1 |
                                 |  1  0 -1 -2 |
o9 : LinearCode
i7 : C=RSCode(GF 5,{1,2,3},3)
o7 = Code with Generator Matrix: | 1  1  1 |
                                 | 1  2 -2 |
                                 | 1 -1 -1 |
o7 : EvaluationCode
\end{verbatim}
\end{example}
\section{Applications of linear codes}
A basic application of a linear code is {\it decoding}, which is used for reliable transmission of information trough a noisy channel. In a few words the idea is the following. Take a vector $c\in C.$ Change the value of some of the entries of $c$ to obtain a new vector $v$. Decoding the vector $v$ means to recover the vector $c$ when only $v$ and $C$ are given. Detailed treatment of decoding algorithms can be found in \cite{huf-pless}. Another, more recent application of linear codes for found in distributed and cloud storage systems. The idea is to use {\it locally recoverable codes}, which are linear codes with the property that every entry can be recovered from a few other entries. For more information on locally recoverable codes see~\cite{TamoBarg}.

Some of the most important functions from this package that can be used for applications of coding theory are the following:
\begin{itemize}
\item syndromeDecode(C,$v$,minimumWeight(C))
\item LocallyRecoverableCode(List,List,a polynomial)
\end{itemize}

Here is a small example.

\begin{example}
$\,$
\begin{verbatim}
i2 : C = HammingCode(2,3);
i3 : msg = matrix {{1,0,1,0}};
i4 : v = msg*(C.GeneratorMatrix)
o4 = | 0 1 0 1 0 1 0 |
i5 : err = matrix take(random entries basis source v, 1)
o5 = | 0 0 0 0 1 0 0 |
i6 : received = transpose(transpose (v+err))
o6 = | 0 1 0 1 1 1 0 |
i7 : transpose syndromeDecode(C, transpose recieved, 3)
o7 = | 0 1 0 1 0 1 0 |
\end{verbatim}
\end{example}

\section*{Acknowledgments}
We thank Federico Galetto, Courtney Gibbons, Hiram L\'opez, and Branden Stone for organizing the Macualay2 workshop at Cleveland State University, where this collaboration started. We want to give a special thanks to Branden Stone for helping us to develop the package during the workshop.

\begin{thebibliography}{99}

\bibitem{carvalho4} C. Carvalho, V. G. Neumann and H. H. L\'opez,
Projective Nested Cartesian Codes,
Bull Braz Math Soc, New Series (2016).

\bibitem{CSTVV} S. M. Cooper, A. Seceleanu, S. O. Toh\v{a}neanu,
M. Vaz Pinto and R. H. Villarreal, 
Generalized minimum distance functions and algebraic invariants of
Geramita ideals, Adv. in Appl. Math. {\bf 112} (2020), 101940.

\bibitem{CLO}D. Cox, J. Little and D. O-Shea,
{\it Ideals, Varieties, and Algorithms}, Springer-Verlag, 1992.

\bibitem{GLS} L. Gold, J. Little, H. Schenck,
{Cayley-Bacharach and evaluation codes on complete intersections},
 J. Pure Appl. Algebra  {\bf 196} (1) (2005) 91--99.

\bibitem{M2} D. R. Grayson and M. E. Stillman,
Macaulay2, a software system for research in algebraic geometry.
Available at \url{http://www.math.uiuc.edu/Macaulay2/}.

\bibitem{Ha1} J. Hansen, { Toric Surfaces and Error--correcting
    Codes} in Coding Theory, Cryptography, and Related Areas, Springer
 (2000), pp. 132-142.

\bibitem{huf-pless} W. Huffman and V. Pless,
{\it Fundamentals of Error-Correcting Codes},
Cambridge University Press, Cambridge, 2003.

\bibitem{LSc} J. Little, H. Schenck,
{\em Toric surface codes and Minkowski sums},
SIAM J. Discrete Math. {\bf  20}  (2006),  no. 4, 999--1014 (electronic).

\bibitem{lopez-villa} H. H. L\'opez, C. Renter\'ia-M\'arquez and R. H. Villarreal,
Affine Cartesian codes,
Des.\ Codes Cryptogr. \textbf{71}(1) (2014) 5--19.

\bibitem{MacWilliams-Sloane} F. J. MacWilliams and N. J. A. Sloane, 
{\it The Theory of Error-correcting Codes}, North-Holland, 1977.

\bibitem{MPV} J. Mart\'inez-Bernal, Y. Pitones and R. H. Villarreal,
Minimum distance functions of graded ideals and Reed-Muller-type codes,
J. Pure Appl. Algebra 221 (2017), 251--275.

\bibitem{MPV2} J. Mart\'inez-Bernal, Y. Pitones and R. H. Villarreal,
Minimum distance functions of complete intersections,
J. Algebra Appl. 17 (2018), no. 11, 1850204 (22 pages).

\bibitem{algcodes} C. Renter\'\i a, A. Simis and R. Villarreal,
Algebraic methods for parameterized codes and invariants of vanishing ideals over finite fields,
Finite Fields Appl. {\bf 17} (2011), no. 1, 81--104.

\bibitem{Ru}D. Ruano, 
{On the parameters of $r$-dimensional toric codes},
Finite Fields and Their Applications {\bf 13} (2007), pp. 962--976.

\bibitem{Shannon} C. Shannon,
A Mathematical Theory of Communication,
The Bell System Technical Journal, {\bf 27} (1948), no. 3, 379--423.

\bibitem{SoSo} I.~Soprunov, J.~Soprunova, {Toric surface codes and Minkowski length of polygons}, SIAM J. Discrete Math. {\bf 23}, Issue 1, (2009) 384-400

\bibitem{So1} I. Soprunov,  {Toric complete intersection codes.} Journal of Symbolic Computation. {\bf 50}, (2013) 374--385.

\bibitem{TamoBarg} I. Tamo and A. Barg,
A Family of Optimal Locally Recoverable Codes,
IEEE Transactions on Information Theory {\bf 60} (2014), no. 8, 4661--4676.

\bibitem{renteria-tapia-ca2} C. Renter\'\i a and H. Tapia-Recillas, 
Reed-Muller codes: an ideal theory approach, Comm. Algebra {\bf 25}
(1997), no. 2, 401--413.

\bibitem{van-lint} J. H. Van Lint,
{ Introduction to coding theory},
Third edition, Graduate Texts in Mathematics {\bf 86}, Springer-Verlag, Berlin, 1999.

\bibitem{monalg} R. H. Villarreal, {\it Monomial Algebras\/},
Second edition, 
Monographs and Research Notes in Mathematics, Chapman and Hall/CRC,
Boca Raton, FL, 2015.
\end{thebibliography}

\end{document}

### 2021-05-13 tex/paper.tex
Fixed issue caused by differences between Macaulay2 1.16 and 1.17, fixed a few minor formatting problems, removed tracking from .dvi and .pdf files
May 13

@@
 \begin{document}
-\title{Coding theory package for Macaulay2}
+\title{Coding theory package for {\it Macaulay2}}
 %%%%%%%%%%%%%%%%%%%%%%%
@@
 \begin{abstract}
-In this Macaulay2 \cite{M2} package we define an object called {\it linear code}.
+In this {\it Macaulay2} \cite{M2} package we implement an object called {\tt LinearCode}.
 We implement functions that compute basic parameters and objects associated with a linear code, such as generator and parity check matrices, the dual code, length, dimension, and minimum distance, among others.
-We define an object {\it evaluation code}, a construction which allows to study linear codes using tools of
+We implement an object called {\tt EvaluationCode}, a construction which allows to study linear codes using tools of
 algebraic geometry and commutative algebra.
 We implement functions to generate important families of linear codes such as Hamming codes, cyclic codes, Reed--Solomon codes, Reed--Muller codes, Cartesian codes, monomial--Cartesian codes, and toric codes.
-In addition, we define functions for the syndrome decoding algorithm and locally recoverable code construction, which are important tools in applications of linear codes.
-The package \textit{CodingTheory.m2} is available at \url{https://github.com/Macaulay2/Workshop-2020-Cleveland/tree/CodingTheory/CodingTheory}
+In addition, we implement functions for the syndrome decoding algorithm and locally recoverable code construction, which are important tools in applications of linear codes.
+The package \textit{CodingTheory} is available at \url{https://github.com/Macaulay2/Workshop-2020-Cleveland/tree/CodingTheory/CodingTheory}
 \end{abstract}
@@
 
-In this note we introduce a new package written for Macaulay2 \cite{M2} called \textit{CodingTheory}.
+In this note we introduce the new {\it Macaulay2} \cite{M2} package called {\tt CodingTheory}.
 The goal of this package is to provide a range of functions for constructing  linear and evaluation codes over finite fields, and for computing some of their main properties.
-To this aim, we define two objects, namely \textit{linear code} and \textit{evaluation code}.
+To this aim, we implement two types of objects, {\tt LinearCode} and {\tt EvaluationCode}.
 The package also includes implementation of functions for generating important families of linear codes like Hamming codes, cyclic codes, Reed-Solomon codes, Reed-Muller codes, Cartesian codes, monomial-Cartesian codes and toric codes.
@@
 The organization of this note is as follows.
-In Section~\ref{def} we describe different ways to define a linear code over a finite field using the \textit{CodingTheory} package.
+In Section~\ref{def} we describe various ways to construct a linear code over a finite field using the \textit{CodingTheory} package.
 In Section~\ref{pro} we show how to compute the main parameters of a linear code: length, dimension, and minimum distance.
@@
 In this paper we do not attempt to fully explain every function distributed in this package.
-%It is important to mention that the functions are not fully explained in this paper. 
-For a detailed explanation of all functions in the package, we refer to the Macaulay2 help page which can be accessed by
+For a detailed explanation of all functions in the package, we refer to the {\it Macaulay2} help page which can be accessed by
 running
-%installPackage("CodingTheory", RemakeAllDocumentation=>true)
+
+\medskip
+
 \begin{verbatim}
-viewHelp CodingTheory
+i1: viewHelp CodingTheory
 \end{verbatim}
+
+\medskip
 
@@
 
-\section{Defining linear codes}\label{def}
+\section{Constructing linear codes}\label{def}
+
 Let $\mathbb{F}_q$ be a finite field with $q$ elements.
-Mathematically, a {\it linear code} is defined as a vector subspace $C\subseteq \mathbb{F}_q^n.$ For Macaulay2 (M2), a linear code is a submodule of $\mathbb{F}_q^n.$ Assume $q=p^r,$ where $p$ is a prime number and $r$ a positive integer.
-By definition, the dual code $C^\perp$ is the orthogonal complement of $C$ in $\mathbb{F}_q^n$ with respect to the standard inner product.
-One can define $C$ by specifying a list $L$ of elements of $\mathbb{F}_q^n$ that span $C$ or by giving a
-a {\it generator matrix} $G$ whose rows form a basis of $C$.
-Alternatively, one can specify a list $L_{H}$ of elements of $\mathbb{F}_q^n$ that span the dual  code $C^\perp$ or a {\it parity check matrix} $H$ whose columns form a basis of the dual code $C^\perp$.
-Below are the commands for the constructor linearCode to construct these equivalent instances of the LinearCode type:
-\begin{itemize}
-\item linearCode($\mathbb{F}_q$,$L$)
-%\item linearCode($\mathbb{F}_q^n$, $L$)
-%\item linearCode($\mathbb{F}_q^n$, $L_H$, ParityCheck $=>$ true)
-%\item linearCode($\mathbb{F}_q$,$L_{H}$, ParityCheck $=>$ true)
-\item linearCode($\mathbb{F}_q$,$n$,$L$)
-\item linearCode($G$)
-\item linearCode($\mathbb{F}_q$,$L_H$, ParityCheck $=>$ true)
-\item linearCode($H$, ParityCheck $=>$ true)
-\item linearCode($p$,$r$,$n$,$L$)
-\item linearCode($p$,$r$,$n$,$L_H$, ParityCheck $=>$ true)
-\end{itemize}
-
-Now, here is a more specific example of how to construct a simple linear code:
-
-\begin{example}
-$\,$
-\begin{verbatim}
-i2 : F = GF 4;
-i3 : L = {{1,1,0,0},{0,0,1,1}};
-i4 : C = linearCode(F,L)
-o4 = Code with Generator Matrix: | 1 1 0 0 |
-                                 | 0 0 1 1 |
-o4 : LinearCode
-\end{verbatim}
-\end{example}
-
-One way to refer to a primitive element of a finite field is by specifying a symbol using the Variable option of the constructor GF.
-
-\begin{example}
-$\,$
-\begin{verbatim}
-i2 : F = GF(9,Variable => a);
-i3 : LH = {{1,0,a,0,0},{0,a,a+1,1,0},{1,1,1,a,0}};
-i4 : C = linearCode(F,LH,ParityCheck => true)
-o4 = Code with Generator Matrix: | a-1 0 a+1 1 0 |
-                                 | 0   0 0   0 1 |
-o4 : LinearCode
-\end{verbatim}
-\end{example}
-To construct a linear code from a matrix, it is necessary to correctly specify the underlying field. This can be done by passing a field to the matrix constructor.
-\begin{example}
-$\,$
-\begin{verbatim}
-i2 : F = GF 4;
-
-i3 : M = matrix(F, {{1,0,1,0},{0,1,1,1}});
-
-             2       4
-o3 : Matrix F  <--- F
-
-i4 : C = linearCode(M)
-
-o4 = Code with Generator Matrix: | 1 0 1 0 |
-                                 | 0 1 1 1 |
-
-o4 : LinearCode
-
-\end{verbatim}
-\end{example}
-
-\section{Basic parameters linear codes}~\label{pro}
-
-The {\it dimension\/} and the {\it length\/} are two of the basic parameters of a code $C\subseteq \mathbb{F}_q^n$.
-They are defined as the subspace dimension $\dim_{\mathbb{F}_q} C$ and the ambient space dimension $n$, respectively.
-A third basic parameter is the {\it minimum weight}, which is given by
-\[
-\min\{\|c\|
-\colon  c \in C, c \neq 0\},
-\]
-where $\| c \|$ is the number of non-zero entries of $c$.
-The {\it rate} of $C$ is defined as the rational number $k/n.$ %The dual of $C$ is defined as the orthogonal complement of $C$ in $\mathbb{F}^n_q$ with respect to the usual dot product in $\mathbb{F}^n_q$.
- Some of the functions that can be used in M2 to compute basic parameters and algebraic objects associated with linear codes are the following: \newline
-\begin{minipage}[t]{0.33\textwidth}
-\begin{itemize}[leftmargin=1 cm]
-\item C.GeneratorMatrix
-\item C.Generators
-\item C.ParityCheckMatrix
-\item C.AmbientModule
-\item alphabet C
-\end{itemize}
-\end{minipage}
-\begin{minipage}[t]{0.33\textwidth}
-\begin{itemize}
-\item field C
-\item informationRate C
-\item ambientSpace C
-\item length C
-\item dim C
-\end{itemize}
-\end{minipage}
-\begin{minipage}[t]{0.33\textwidth}
-\begin{itemize}
-\item minimumWeight C
-\item codewords C
-\item dualCode C
-\item shorten(C, List)
-\item ==
-\end{itemize}
-\end{minipage}
-
-\begin{example}
-$\,$
-\begin{verbatim}
-i2: F = GF 4;
-i3 : L = {{1,1,0,0},{0,0,1,1}};
-i4 : C = linearCode(F,L)
-o4 = Code with Generator Matrix: | 1 1 0 0 |
-                                 | 0 0 1 1 |
-o4 : LinearCode
-i5 : length C
-o5 = 4
-i6 : dim C
-o6 = 2
-i7 : informationRate C
-     1
-o7 = -
-     2
-o7 : QQ
-i8 : ambientSpace C
-      4
-o8 = F
-o8 : F-module, free
-i9 : alphabet C
-o9 = {0, a, a + 1, 1}
-o9 : List
-i10 : minimumWeight C
-o10 = 2
-i11 : dualCode C
-o11 = Code with Generator Matrix: | 1 1 0 0 |
-                                  | 0 0 1 1 |
-
-o11 : LinearCode
-\end{verbatim}
-\end{example}
-
-\section{Evaluation codes}\label{eva}
-Let $\mathcal{X}=\left\{{\bf a}_1,\ldots,{\bf a}_n\right\}$ be a subset of an $m$-dimensional space $\mathbb{F}_q^{m}$.
-Consider a finite dimensional subspace $S\subset {\mathbb{F}_q}[X_1,\ldots,X_m]$  of the $m$-variate polynomial ring over $\mathbb{F}_q$.
-%and $S_{\leq d}$ the ${\mathbb{F}_q}$-vector space of all polynomials in $S$ of degree at most $d.$
-
-The {\it evaluation map\/}
-\begin{equation*}\label{functionevaluation}
-{\rm ev}_S\colon S\longrightarrow {\mathbb{F}_q}^{|\mathcal{X}|},\ \ \ \ \ 
-f\mapsto \left(f({\bf a}_1),\ldots,f({\bf a}_n)\right),
-\end{equation*}
-defines a linear map of ${\mathbb{F}_q}$-vector spaces.
-The image of ${\rm ev}_S$ in ${\mathbb{F}_q}^{|\mathcal{X}|}$, denoted by $C_{\mathcal{X}}(S)$, %defines a ${\mathbb{F}_q}$-vector subspace.
-We refer to $C_{\mathcal{X}}(d)$
-is the {\it evaluation code\/} on the set $\mathcal{X}$ corresponding to $S$.
-The {\it vanishing ideal} of $\mathcal{X}$, denoted by $I(\mathcal{X})$, is the ideal in $S$ of all polynomials that vanish on  $\mathcal{X}.$ A key observation that allows the use of commutative algebra in studying evaluation codes is that the kernel of the evaluation map ${\rm ev}_S$ is precisely $S\cap I(\mathcal{X})$.
-
-An evaluation code $C_{\mathcal{X}}(S)$ is defined in M2 as a separate type because there are more objects associated with it than with a linear code.
-For instance, the vanishing ideal associated to the set $\mathcal{X}$ plays an important role when finding and estimating parameters of the code, so it is convenient to be able to access it.
-Given an evaluation code $C$ in M2, the object C.linearCode is a linear code in M2.
-The command $\text{evaluationCode}(\mathbb{F}_q,\mathcal{X},L)$ defines an evaluation code where $\mathcal{X}$ is a list of elements in $\mathbb{F}_q^{m}$ and $L$ is a list of polynomials that span $S$.
-In the case when polynomials in $L$ are monomials, one may give the matrix of exponent vectors instead of $L$.
-
-There are many construction of evaluation codes for specific choices of the set $\mathcal{X}$ and the subspace $S$. These include
-Reed-Muller codes, Cartesian and monomial Cartesian codes, toric codes, and evaluation codes from graphs. We refer to 
-\cite{carvalho4, Ha1, LSc, lopez-villa, MPV, algcodes, Ru, SoSo} for details on how these codes are defined and what properties they have from coding theory, commutative algebra, and algebraic geometry perspectives.
-Some functions defined in this package for various constructions of evaluation codes and associated algebraic objects are the following:
+Mathematically, a {\it linear code} is defined as a vector subspace $C\subseteq \mathbb{F}_q^n$ and it is often specified by a {\it generator matrix}, which is a $k \times n$ matrix $G$ with entries in $\mathbb{F}_q$ whose $k$ rows form a basis for $C$.
+In {\it Macaulay2}, a linear code is defined as an $\mathbb{F}_q$-submodule of $\mathbb{F}_q^n$ using the constructor {\tt linearCode}.
+This constructor is an instance of the {\tt LinearCode} type.
+There are various ways to use the command {\tt linearCode}.
+For example, one can use this command to construct a linear code $C\subseteq \mathbb{F}^n_q$ by specifying a generator matrix $G$ of $C$ (Example \ref{example_2.1}).
+Alternatively, one can use the command {\tt linearCode} to construct a linear code $C\subseteq \mathbb{F}_q^n$ by indicating the finite field $\mathbb{F}_q$ and a list $L$ of elements of $\mathbb{F}_q^n$ that span $C$ (Example \ref{example_2.2}).
+More details and equivalent ways to use the constructor {\tt linearCode} are given next.
 
@@
 
-\begin{minipage}[t]{0.52\textwidth}
+\begin{minipage}[t]{0.6\textwidth}
+ Inputs:
  \begin{itemize}[leftmargin=0.55 cm]
-\item evaluationCode($\mathbb{F}_q$,List,List)
-\item toricCode($\mathbb{F}_q$,Integer matrix)
-\item cartesianCode($\mathbb{F}_q$,List,List)
-\item orderCode($\mathbb{F}_q$,List,List,ZZ)
-\item evCodeGraph($\mathbb{F}_q$,Incident Matrix,integer)
+  \item {\tt F = GF(q)}, a finite field with {\tt q} elements
+  \item {\tt n, r, p}, positive integers with {\tt p} prime
+  \item {\tt G}, a matrix with entries in {\tt GF(q)}
+  \item {\tt L}, a list of elements of $\mathtt{GF(q)}^\mathtt{n}$
  \end{itemize}
 \end{minipage}
-\begin{minipage}[t]{0.45\textwidth}
-\begin{itemize}[leftmargin=1 cm]
-\item vNumber(Ideal)
-\item footPrint(Integer,Integer,Ideal)
-\item hYpFunction(Integer,Integer,Ideal)
-\item gMdFunction(Integer,Integer,Ideal)
-\item vasFunction(Integer,Integer,Ideal)
+\begin{minipage}[t]{0.35\textwidth}
+Usage:
+\begin{itemize}[leftmargin=0.55 cm]
+\item {\tt linearCode(G)}
+\item {\tt linearCode(F,L)}
+\item {\tt linearCode(F,n,L)}
+\item {\tt linearCode(p,r,n,L)}
 \end{itemize}
@@
 
-The mathematical definitions of the vNumber, the footprint function, the hyp function, the generalized footprint function and the Vasconcelos function can be found in~\cite{CSTVV}.
-The following example shows how to construct an evaluation code.
+In the next examples we construct a simple binary linear code $C\subseteq \mathbb{F}^4_2$ with generator matrix 
+$G=
+\begin{pmatrix}
+ 1 & 1 & 0 & 0\\
+ 0 & 0 & 1 & 1\\
+\end{pmatrix}
+$ using equivalent versions of the command {\tt linearCode}.
+
+\begin{example}\label{example_2.1}
+ $\,$
+\begin{verbatim}
+i1 : F = GF(2);
+i2 : L = {{1,1,0,0},{0,0,1,1}};
+i3 : G = matrix apply(L,codeword->apply(codeword,entry->sub(entry,F)));
+i4 : C = linearCode(G);
+i5 : C.GeneratorMatrix
+o5 = | 1 1 0 0 |
+     | 0 0 1 1 |
+ \end{verbatim}
+\end{example}
+\vskip -0.5 cm
+Note that in Example \ref{example_2.1} was necessary to coerce the entries of each vector in the list {\tt L} into elements of {\tt F = GF(2)}.
+An equivalent way to do this is to pass the field {\tt GF(q)} to the {\tt matrix} constructor or to use the constructor {\tt linearCode(GF(q),L)}.
+
+\begin{example}\label{example_2.2}
+ $\,$
+\begin{verbatim}
+i1 : L = {{1,1,0,0},{0,0,1,1}};
+i2 : C = linearCode(GF(2),L);
+i3 : C.GeneratorMatrix
+o3 = | 1 1 0 0 |
+     | 0 0 1 1 |
+\end{verbatim}
+\end{example}
+The set $\mathbb{F}^{\ast}_q$ of non-zero elements of a finite field $\mathbb{F}_q$ is a multiplicative cyclic group (\cite[Theorem 3.3.1]{huf-pless}).
+A generator of $\mathbb{F}_q^{\ast}$ is called a \textit{primitive element} of $\mathbb{F}_q$.
+One way to refer to a primitive element of a finite field is by specifying a symbol using the {\tt Variable} option of the constructor {\tt GF}.
+In the next example we illustrate how to define a linear code  $C\subseteq \mathbb{F}_{11}^{10}$ with generator matrix
+\[
+ G=
+ \begin{pmatrix}
+  1 & a^1 & a^2 & \cdots & a^9\\
+  1 & a^2 & a^4 & \cdots & (a^2)^9\\
+  1 & a^3 & (a^3)^2 & \cdots & (a^3)^9\\
+  1 & a^4 & (a^4)^2 & \cdots & (a^4)^9\\
+ \end{pmatrix},
+\]
+where $a$ is a primitive element of $\mathbb{F}_{11}$. In {\it Macaulay2}, $a=2$.
+
+\begin{example}\label{example_2.4}
+ $\,$
+\begin{verbatim}
+i1 : F = GF(11,Variable => a);
+i2 : G = matrix table({1,2,3,4},
+                      {1,a,a^2,a^3,a^4,a^5,a^6,a^7,a^8,a^9},(i,j)->j^i);
+i3 : C = linearCode(G);
+i4 : C.GeneratorMatrix
+o4 = | 1 2  4  -3 5  -1 -2 -4 3  -5 |
+     | 1 4  5  -2 3  1  4  5  -2 3  |
+     | 1 -3 -2 -5 4  -1 3  2  5  -4 |
+     | 1 5  3  4  -2 1  5  3  4  -2 |
+\end{verbatim}
+\end{example}
+
+In $\mathbb{F}_q^n$ there is a standard inner product defined for all $x=(x_1,\ldots,x_n)$ and $y=(y_1,\ldots,y_n)$ in $\mathbb{F}_q^n$ as $x\cdot y=x_1y_1+\cdots+x_ny_n.$ Given a linear code $C\subseteq \mathbb{F}^n_q$, the {\it dual} or {\it orthogonal code} of $C$ with respect to this product is the linear code $C^{\perp}$ defined as
+\[
+C^\perp=\{x\in \mathbb{F}_q^n\,:\,x\cdot c=0 \text{ for all }c\in C\}.
+\]
+A generator matrix of $C^{\perp}$ is called a {\it parity check matrix} for $C$.
+Since $(C^{\perp})^{\perp}=C$, another common way to mathematically define $C$ is by specifying one of its parity check matrices.
+By using the command {\tt linearCode} and the {\tt ParityCheck} option of this command, we can define a linear code $C\subseteq \mathbb{F}^n_q$ by either specifying a parity check matrix $H$ of $C$ or a list $L$ of elements of $\mathbb{F}_q^n$ that span $C^{\perp}$.
+
+%We now present some examples to illustrate how to use the {\tt ParityCheck} option of the command {\tt linearCode}. 
+In the next example, we define a linear code $C\subseteq \mathbb{F}_9^5$ whose dual code $C^{\perp}$ is generated by the set $\{(1,0,a,0,0),(0,1,a+1,1,0),(1,1,1,a,0)\}\subseteq \mathbb{F}^5_9$, where $a$ is a primitive element of $\mathbb{F}_9$.
+In {\it Macaulay2}, $a\in\mathbb{F}_9$ satisfies $a^2=a+1$.
+
+\begin{example}\label{example_2.5}
+$\,$
+\begin{verbatim}
+i1 : F = GF(9,Variable => a);
+i2 : L = {{1,0,a,0,0},{0,a,a+1,1,0},{1,1,1,a,0}};
+i3 : C = linearCode(F,L,ParityCheck => true);
+i4 : C.GeneratorMatrix
+o4 = | a-1 0 a+1 1 0 |
+     | 0   0 0   0 1 |
+i5 : C.ParityCheckMatrix
+o5 = | 1 0 a   0 0 |
+     | 0 a a+1 1 0 |
+     | 1 1 1   a 0 |
+\end{verbatim}
+\end{example}
+
+Although the dual code of a linear code can be constructed using the command {\tt dualCode} implemented in the {\it CodingTheory} package, the {\tt ParityCheck} option of the command {\tt linearCode} also allow us to construct the dual of a linear code.
+In the next example we construct the dual of the linear code in Example \ref{example_2.4}.
+
+\begin{example}\label{example_2.5}
+ $\,$
+ \begin{verbatim}
+i1 : F = GF(11,Variable => a);
+i2 : G = matrix table({1,2,3,4},
+                      {1,a,a^2,a^3,a^4,a^5,a^6,a^7,a^8,a^9},(i,j)->j^i);
+i3 : D = linearCode(G,ParityCheck => true);
+i4 : D.GeneratorMatrix
+o4 = | 1  -3 5  3  1 0 0 0 0 0 |
+     | -3 -1 4  -4 0 1 0 0 0 0 |
+     | 4  -4 -3 5  0 0 1 0 0 0 |
+     | -5 -3 4  4  0 0 0 1 0 0 |
+     | -4 -4 -1 3  0 0 0 0 1 0 |
+     | -3 5  3  1  0 0 0 0 0 1 |
+ \end{verbatim}
+\end{example}
+
+\section{Basic parameters of linear codes}~\label{pro}
+
+The {\it dimension\/}, {\it length\/}, and {\it rate} or {\it information rate}  of a linear code are basic parameters.
+Given a linear code $C\subseteq \mathbb{F}^n_q$, these parameters are defined as $k=\dim_{\mathbb{F}_q} C$, $n$, and $k/n$ respectively.
+Another parameter of a linear code $C\subseteq \mathbb{F}^n_q$ is the {\it minimum Hamming weight}, which is given by
+\[
+w_H(C)=\min\{\|c\|
+\colon  c \in C, c \neq 0\},
+\]
+where $\| c \|$ is the number of non-zero entries of $c\in \mathbb{F}^n_q$.
+This parameter is important in determining the error-correcting capability of $C$; the higher the minimum Hamming weight, the more errors the code can detect and correct (see \cite{huf-pless}).
+This function has been implemented \textcolor{red}{Nathan, write here something about the implementation of the weight function}.
+
+In {\it Macaulay2}, the space $\mathbb{F}_q^n$ has been called the {\it ambient module} or {\it ambient space} of $C$, while the field $\mathbb{F}_q$, the {\it alphabet} or {\it field} of $C$.
+As often, the elements of $C$ are referred to as {\it codewords}.
+The input in all the following commands implemented in the {\tt CodingTheory} package is always a linear code  {\tt C}.
+
+\medskip
+
+\begin{minipage}[t]{0.3\textwidth}
+\begin{itemize}[leftmargin=0.5 cm]
+\item {\tt dim C}
+\item {\tt length C}
+\item {\tt alphabet C}
+\item {\tt ambientSpace C}
+\end{itemize}
+\end{minipage}
+\begin{minipage}[t]{0.32\textwidth}
+\begin{itemize}[leftmargin=0.35 cm]
+\item {\tt field C}
+\item {\tt codewords C}
+\item {\tt C.Generators}
+\item {\tt C.GeneratorMatrix}
+\end{itemize}
+\end{minipage}
+\begin{minipage}[t]{0.35\textwidth}
+\begin{itemize}[leftmargin=0.55 cm]
+\item {\tt C.AmbientModule}
+\item {\tt minimumWeight C}
+\item {\tt informationRate C}
+\item {\tt C.ParityCheckMatrix}
+\end{itemize}
+\end{minipage}
+\newpage
+\begin{example}
+$\,$
+\begin{verbatim}
+i1 : L = {{1,1,0,0},{0,0,1,1}};
+i2 : C = linearCode(GF(4),L);
+i3 : dim C
+o3 : 2
+i4 : length C
+o4 = 4
+i5 : alphabet C
+o5 = {0, a, a + 1, 1}
+i6 : ambientSpace C
+           4
+o6 = (GF 4)
+i7 : field C
+o7 = GF 4
+i8 : minimumWeight C
+o8 = 2
+i9 : codewords C
+o9 = {{1, 1, 1, 1}, {1, 1, a, a}, {a, a, 1, 1}, {a, a, a, a}, 
+     ------------------------------------------------------------------
+      {a + 1, a + 1, a, a}, {a + 1, a + 1, 1, 1}, {1, 1, a + 1, a + 1},
+     ------------------------------------------------------------------
+      {a, a, a + 1, a + 1}, {a + 1, a + 1, a + 1, a + 1}, {1, 1, 0, 0},
+     ------------------------------------------------------------------
+      {0, 0, 1, 1}, {0, 0, a, a}, {a, a, 0, 0}, {a + 1, a + 1, 0, 0}, 
+     ------------------------------------------------------------------
+      {0, 0, a + 1, a + 1}, {0, 0, 0, 0}}
+\end{verbatim}
+\end{example}
+\section{Evaluation codes}\label{eva}
+Let $P=\left\{{\bf a}_1,\ldots,{\bf a}_n\right\}$ be a subset of $\mathbb{F}_q^{m}$.
+Consider a finite dimensional subspace $\mathcal{S}\subset {\mathbb{F}_q}[X_1,\ldots,X_m]$  of the ring of polynomials over $\mathbb{F}_q$ in $m$ variables. The {\it evaluation map\/}
+\begin{equation*}\label{functionevaluation}
+{\rm ev}_\mathcal{S}\colon \mathcal{S}\longrightarrow {\mathbb{F}_q}^{|P|},\ \ \ \ \ 
+f\mapsto \left(f({\bf a}_1),\ldots,f({\bf a}_n)\right),
+\end{equation*}
+defines a linear map of ${\mathbb{F}_q}$-vector spaces.
+The image of ${\rm ev}_\mathcal{S}$ in ${\mathbb{F}_q}^{|P|}$, denoted by $C_{P}(\mathcal{S})$, %defines a ${\mathbb{F}_q}$-vector subspace.
+We refer to $C_{P}(d)$
+is the {\it evaluation code\/} on the set $P$ corresponding to $\mathcal{S}$.
+The {\it vanishing ideal} of $P$, denoted by $I(P)$, is the ideal in $\mathbb{F}_q[X_1,\ldots,X_n]$ of all polynomials that vanish on  $P$.
+A key observation that allows the use of commutative algebra in studying evaluation codes is that the kernel of the evaluation map ${\rm ev}_S$ is precisely $\mathcal{S}\cap I(P)$.
+
+An evaluation code $C_{P}(\mathcal{S})$ is implemented in {\it Macaulay2} as an {\tt EvaluationCode}.
+Although, the object {\tt C.LinearCode} is a linear code in {\it Macaulay2}.
+This has been done in this way because there are more objects associated with an evaluation code than with a linear code.
+For instance, the vanishing ideal associated to the set $P$ plays an important role when finding and estimating parameters of the code, so it is convenient to be able to access it.
+
+There are many constructions of evaluation codes for specific choices of the set $P$ and the subspace $\mathcal{S}$. These include
+Reed-Muller codes, Cartesian, monomial Cartesian codes, toric codes, and evaluation codes from graphs. We refer to 
+\cite{carvalho4, Ha1, LSc, lopez-villa, MPV, algcodes, Ru, SoSo} for details on how these codes are defined and what properties they have from coding theory, commutative algebra, and algebraic geometry perspectives.
+
+Some functions implemented in the {\it CodingTheory} package for various constructions of evaluation codes and associated algebraic objects are the following:
+
+\medskip
+
+Inputs:
+
+\medskip
+
+\begin{minipage}[t]{0.53\textwidth}
+\begin{itemize}[leftmargin=0.55 cm]
+  \item {\tt I}, an ideal
+  \item {\tt d,r}, positive integers
+  \item {\tt F = GF(q)}, a finite field with {\tt q} elements
+  \item {\tt P}, a list of points in $\mathtt{F}^\mathtt{m}$
+  \item {\tt S}, a list of polynomials in {\tt m} variables
+\end{itemize}
+\end{minipage}
+\begin{minipage}[t]{0.45\textwidth}
+\begin{itemize}[leftmargin=0.55 cm]
+  \item {\tt M}, an integer matrix
+  \item {\tt L}, a list of subsets of {\tt F}
+  \item {\tt v}, a list of {\tt m} positive integers
+  \item {\tt MI}, an incident matrix of a graph
+ \end{itemize}
+\end{minipage}
+
+\medskip
+
+Usage:
+
+\medskip
+
+\begin{minipage}[t]{0.53\textwidth}
+\begin{itemize}[leftmargin=0.55 cm]
+\item {\tt evaluationCode(F,P,S)}
+\item {\tt toricCode(F,M)}
+\item {\tt cartesianCode(F,L,d)}
+\item {\tt orderCode(F,P,v,d)}
+\item {\tt evCodeGraph(F,MI,S)}
+\end{itemize}
+\end{minipage}
+\begin{minipage}[t]{0.45\textwidth}
+\begin{itemize}[leftmargin=0.55 cm]
+\item {\tt vNumber(I)}
+\item {\tt footPrint(d,r,I)}
+\item {\tt hYp(d,r,I)}
+\item {\tt genMinDisIdeal(d,r,I)}
+\item {\tt vasconcelosDegree(d,r,I)}
+\end{itemize}
+\end{minipage}
+
+\medskip
+
+The mathematical definitions of the functions in the second column of the previous list can be found in \cite{CSTVV}.
+The following example shows how to construct an evaluation code in {\it Macaulay2} using the {\it CodingTheory} package.
 
@@
 \begin{verbatim}
-i2 : F=GF(4); R=F[x,y,z];
-i4 : P={{0,0,0},{1,0,0},{0,1,0},{0,0,1},{1,1,1},{a,a,a}};
-i5 : S={x+y+z,a+y*z^2,z^2,x+y+z+z^2};
-i6 : C=evaluationCode(F,P,S)
-o6 = Code with Generator Matrix: | 0 1 1 1 1   a   |
+i1 : F=GF(4,Variable=>a); R=F[x,y,z];
+i3 : P={{0,0,0},{1,0,0},{0,1,0},{0,0,1},{1,1,1},{a,a,a}};
+i4 : S={x+y+z,a+y*z^2,z^2,x+y+z+z^2};
+i5 : C=evaluationCode(F,P,S);
+i6 : (C.LinearCode).GeneratorMatrix
+o6 = | 0 1 1 1 1   a   |
      | a a a a a+1 a+1 |
@@
      | 0 1 1 0 0   1   |
-o6 : EvaluationCode
+
 i7 : length C.LinearCode
@@
 o9 = {{0, 0, 0}, {1, 0, 0}, {0, 1, 0}, {0, 0, 1}, {1, 1, 1}, {a, a, a}}
-o9 : List
-i10 : C.VanishingIdeal;
-o10 = Ideal of R
+
+i10 : C.VanishingIdeal
+
+                         2    2                      2    2              
+o10 = ideal (x*z + y*z, y  + z  + y + z, x*y + y*z, x  + z  + x + z, 
+      --------------------------------------------------------------
+      3           2
+     z  + (a +1)z  + a*z)
 \end{verbatim}
 \end{example}
+
 \section{Families of linear codes}\label{families}
-We continue with the same notation: $n$ represents the length of the code, $k$ the dimension and $q$ the size of the field.
-Some families of linear codes that have been implemented in this package are the following:
+
+Various important families of linear codes have been defined through the development of coding theory.
+These include Hamming codes, cyclic codes, Reed-Solomon codes, Reed-Muller codes, etc.
+The mathematical definitions of these and many other families of codes can be found in \cite{huf-pless,MacWilliams-Sloane,van-lint}.
+
+The following functions have implemented in the {\it CodingTheory} package to construct some of these important families of codes.
 
@@
 
-\begin{minipage}[t]{0.5\textwidth}
-\begin{itemize}[leftmargin=1 cm]
-\item HammingCode($q$,integer)
-\item randLDPC($n$, $k$, integer, integer)
-\item cyclicCode($\mathbb{F}_q$ ,polynomial, $n$)
-\item quasiCyclicCode($\mathbb{F}_q$,list)
-\item RSCode($\mathbb{F}_q$,List,integer)
-\item RMCode($q$,integer,integer)
+Inputs:
+
+\medskip
+
+\begin{minipage}[t]{0.53\textwidth}
+\begin{itemize}[leftmargin=0.55 cm]
+  \item {\tt n, k, d, m, s}, positive integers
+  \item {\tt F = GF(q)}, a finite field with {\tt q} elements
+  \item {\tt g(x)}, a polynomial in {\tt F[x]}
  \end{itemize}
 \end{minipage}
-\begin{minipage}[t]{0.5\textwidth}
-\begin{itemize}[leftmargin=1 cm]
-\item zeroCode($\mathbb{F}_q$,$n$)
-\item universeCode($\mathbb{F}_q$,$n$)
-\item repetitionCode($\mathbb{F}_q$,$n$)
-\item zeroSumCode($\mathbb{F}_q$,$n$)
-\item random($\mathbb{F}_q$, $n$, $k$)
-
+\begin{minipage}[t]{0.45\textwidth}
+ \begin{itemize}[leftmargin=0.55 cm]
+  \item {\tt E}, a list of elements of {\tt F}
+  \item {\tt L}, a list of vectors in $F^{\mathtt{m}}$
  \end{itemize}
@@
 
-Mathematical definitions of the above families can be found in \cite{huf-pless,MacWilliams-Sloane,van-lint}.
+Usage:
+
+\medskip
+
+\begin{minipage}[t]{0.35\textwidth}
+\begin{itemize}[leftmargin=0.55 cm]
+\item {\tt hammingCode(q,s)}
+\item {\tt reedSolomonCode(F,E,s)}
+\item {\tt reedMullerCode(q,m,d)}
+\end{itemize}
+\end{minipage}
+\begin{minipage}[t]{0.33\textwidth}
+\begin{itemize}[leftmargin=0.55 cm]
+\item {\tt cyclicCode(F,g(x),n)}
+\item {\tt repetitionCode(F,n)}
+\item {\tt randomCode(F,n,k)}
+%\item {\tt quasiCyclicCode(F,L)}
+\end{itemize}
+\end{minipage}
+\begin{minipage}[t]{0.3\textwidth}
+\begin{itemize}[leftmargin=0.55 cm]
+\item {\tt zeroCode(F,n)}
+\item {\tt universeCode(F,n)}
+\item {\tt zeroSumCode(F,n)}
+\end{itemize}
+\end{minipage}
+
+\medskip
 
@@
 \begin{verbatim}
-i2 : C = HammingCode(2,3)
-o2 = Code with Generator Matrix: | 1 1 1 1 0 0 0 |
-                                 | 0 1 1 0 1 0 0 |
-                                 | 1 0 1 0 0 1 0 |
-                                 | 1 1 0 0 0 0 1 |
-o2 : LinearCode
-i3 : F=GF(5); R=F[x]; g=x-1; C=cyclicCode(F,g,6)
-Cyclic Code
-o6 = Code with Generator Matrix: |-1  1  0  0  0  0 |
+i1 : C = hammingCode(2,3);
+i2 : C.GeneratorMatrix
+o2 = | 1 1 1 1 0 0 0 |
+     | 1 1 0 0 1 0 0 |
+     | 0 1 1 0 0 1 0 |
+     | 1 0 1 0 0 0 1 |
+
+i3 : F = GF(5); R = F[x]; g = x-1; C = cyclicCode(F,g,6); 
+i7 : C.GeneratorMatrix
+o7 = | -1 1  0  0  0  0 |
      | 0  -1 1  0  0  0 |
@@
      | 0  0  0  0  -1 1 |
-o6 : LinearCode
-i7 : F = GF(5);
-i8 : L = apply(toList(1..2),j-> apply(toList(1..4),i-> random(F)))
-o8 = {{0, -2, 2, -1}, {-2, 1, 0, -1}}
-o8 : List
-i9 : C=quasiCyclicCode(L)
-o9 = Code with Generator Matrix: |  0 -2  2 -1 |
-                                 | -1  0 -2  2 |
-                                 |  2 -1  0 -2 |
-                                 | -2  2 -1  0 |
-                                 | -2  1  0 -1 |
-                                 | -1 -2  1  0 |
-                                 |  0 -1 -2  1 |
-                                 |  1  0 -1 -2 |
-o9 : LinearCode
-i7 : C=RSCode(GF 5,{1,2,3},3)
-o7 = Code with Generator Matrix: | 1  1  1 |
+
+i8 : C = reedSolomonCode(GF(5),{1,2,3},3);
+i9 : (C.LinearCode).GeneratorMatrix
+o9 = | 1 1  1  |
      | 1 2  -2 |
      | 1 -1 -1 |
-o7 : EvaluationCode
 \end{verbatim}
@@
 \section{Applications of linear codes}
-A basic application of a linear code is {\it decoding}, which is used for reliable transmission of information trough a noisy channel.
+An important aspect in coding theory is {\it decoding}, which is used when information is transmitted trough a noisy channel.
 In a few words the idea is the following.
@@
 Detailed treatment of decoding algorithms can be found in \cite{huf-pless}.
-Another, more recent application of linear codes for found in distributed and cloud storage systems.
+Another, more recent application of coding theory is found in distributed and cloud storage systems.
 The idea is to use {\it locally recoverable codes}, which are linear codes with the property that every entry can be recovered from a few other entries.
-For more information on locally recoverable codes see~\cite{TamoBarg}.
+For more information on locally recoverable codes see \cite{TamoBarg}.
 
-Some of the most important functions from this package that can be used for applications of coding theory are the following:
+Some of the most important functions implemented in the {\it CodingTheory} package that can be used for applications of coding theory are the following:
+
+\medskip
+
+Inputs:
+
+\medskip
+
 \begin{itemize}
-\item syndromeDecode(C,$v$,minimumWeight(C))
-\item LocallyRecoverableCode(List,List,a polynomial)
+ \item {\tt C}, a linear code over {\tt GF(q)}
+ \item {\tt v}, a vector in the ambient space of {\tt C}
+ \item {\tt \{q,n,k,r\}}, where {\tt q} is a prime power, and {\tt n}, {\tt k} and
+      {\tt r} are positive integers
+ \item {\tt L}, a pairwise disjoint list of subsets of elements of {\tt GF(q)}
 \end{itemize}
 
-Here is a small example.
+\medskip
+
+Usage
+
+\medskip
+
+\begin{itemize}
+\item {\tt syndromeDecode(C,$v$,minimumWeight(C))}
+\item {\tt LocallyRecoverableCode(\{q,n,k,r\},L,a polynomial)}
+\end{itemize}
+
+\medskip
 
@@
 \begin{verbatim}
-i2 : C = HammingCode(2,3);
-i3 : msg = matrix {{1,0,1,0}};
-i4 : v = msg*(C.GeneratorMatrix)
-o4 = | 0 1 0 1 0 1 0 |
-i5 : err = matrix take(random entries basis source v, 1)
-o5 = | 0 0 0 0 1 0 0 |
-i6 : received = transpose(transpose (v+err))
-o6 = | 0 1 0 1 1 1 0 |
-i7 : transpose syndromeDecode(C, transpose recieved, 3)
-o7 = | 0 1 0 1 0 1 0 |
+i1 : C = hammingCode(2,3);
+i2 : msg = matrix {{1,0,1,0}};
+i3 : v = msg*(C.GeneratorMatrix)
+o3 = | 0 1 0 1 0 1 0 |
+i4 : err = matrix take(random entries basis source v, 1)
+o4 = | 0 0 0 0 1 0 0 |
+i5 : received = transpose(transpose (v+err))
+o5 = | 0 1 0 1 1 1 0 |
+i6 : transpose syndromeDecode(C, transpose recieved, 3)
+o6 = | 0 1 0 1 0 1 0 |
 \end{verbatim}
@@
 We want to give a special thanks to Branden Stone for helping us to develop the package during the workshop.
+This work was partially supported by the NSF grant DMS-2003883.
 
@@
 \bibitem{M2} D. R. Grayson and M. E. Stillman,
-Macaulay2, a software system for research in algebraic geometry.
+{\it Macaulay2}, a software system for research in algebraic geometry.
 Available at \url{http://www.math.uiuc.edu/Macaulay2/}.
@@
 \end{thebibliography}
-
 \end{document}

### 2021-05-13 tex/answers.tex

\documentclass[12pt]{amsart}
\frenchspacing \mathsurround=1pt \emergencystretch=5pt
\tolerance=400

\topmargin = -0.1in \evensidemargin = 0.2in \oddsidemargin = 0.2in
\textheight = 23cm \headheight = 10pt
\textwidth = 15.7cm

\usepackage{amsmath, amsthm, latexsym, amssymb, amsfonts, epsfig, epsf, xcolor, hyperref}
\usepackage{mathtools}
\usepackage{bm,xcolor,tikz,hyperref}
%\usepackage{refcheck}
\usepackage{bm}
\usepackage{blkarray}
\usepackage{booktabs}
\usepackage{enumitem}

\newenvironment{pf}{\proof[\proofname]}{\endproof}
\theoremstyle{plain}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{example}[theorem]{Example}
\newtheorem{remark}[theorem]{Remark}

\newcommand{\rmv}[1]{}

\begin{document}
\begin{center}
Response to reviews
\end{center}
We would like to thank you for your careful reading the manuscript and providing insightful comments. They have served to improve the submission. The first main change is the documentation in the package. We extended/improved/homogenized/latex-it all the documentation. The second main change is the minimum distance function, we found a flaw that \textcolor{red}{Nathan, write something please}. Here is a list of answers to the reviewer's comments that we performed during the preparation of this revision.
\begin{center}
{\bf report\_M\_on\_v1}
\end{center}
\begin{itemize}
\item[\rm 1.] Example 4.1 now contains the output of the command {\tt C.VanishingIdeal}.
\item[\rm 2.] \textcolor{red}{Nathan, please write something about the minimum weight}
\item[\rm 3.] We were not aware of this convention, thank you for pointing it out. We changed the name of the functions.
\item[\rm 7.] The documentation of the function {\tt LocallyRecoverableCode} has been corrected and now has more information.
\item[\rm 9.] In the documentation for shorten, the word shorten in the title has not been capitalized because according to the literature of {\it Macaulay2}, the titles start with a lower letter. Actually, we were not aware of this and now all the titles start with a lower letter.
\item[\rm 10.] The documentation for {\tt tannerGraph} has been extended.
\end{itemize}
We also corrected the typos that are mentioned in items 4, 5, 6, 8, 11.
\begin{center}
{\bf report\_C\_on\_v1} (Code)
\end{center}
\begin{itemize}
\item The package now works with {\it Macaulay2} 1.17.
\item About public domain \textcolor{red}{Gwin, help please}
\item We have updated {\tt PackageImports} and {\tt PackageExports}.
\item About return and semicolon \textcolor{red}{Gwin, help please}
\item We corrected the word Module in the expression ``Expected codewords all to be the same length and equal to the rank of the Module".
\item Now all the functions have their bodies indented.
\end{itemize}
\begin{center}
{\bf report\_C\_on\_v1} (Paper)
\end{center}
\begin{itemize}
\item The word {\it Macaulay2} is always in italic and never abbreviated as M2.
\item We replaced the word ``define'' with ``implement'' where the meaning of the context implies that a function or object has been implemented in this package. For example, In line 1 in the Abstract the phrase ``we define an object called {\it linear code}'' was substituted by ``we implement an object called {\tt LinearCode}.
\item We replaced ``different'' with ``various'' in line 1 in the last paragraph in page 1.
\item We fixed the grammar in ``There are many construction of evaluation codes'' to ``There are many constructions of evaluation codes''.
\item All computer code is in {\tt typewriter} font.
\item We corrected the mixture of computer code and mathematics.
\end{itemize}
\begin{center}
{\bf report\_C\_on\_v1} (Documentation)
\end{center}
\begin{itemize}
\item There are no more warnings about missing documentation. \textcolor{red}{Nathan, there is a missing documentation now about the weight function}
\item We corrected all the typos that you found. And more! We really want to thank you for all the comments, they were the basis to redo the documentation.
\item The headline now is ``tools for Coding Theory".
\item We are using LaTex for all the documentation.
\item HashTable has been changed to hash table. Similar issues have been corrected.
\item We changed ``m-variate polynomials over F'' by ``polynomials over F in m variables''.
\item We corrected the following. This sounds circular:	field -- Returns the field where the entries of the field belong.
Now we have: field -- the field of a code
\item Now we have in the description of ``field" the following: Given a code {\tt C}, returns the field (or ring) that contains the entries of the generator matrix of {\tt C}.
\item We added more information in the description of the following functions: generalized footprint, footprint, v-number, Vasconcelos and hyp. In addition, we added hyperlinks to the papers where they were defined.
\item We have now linearly independent instead of l.i.
\item We are using now: this symbol is used as a key for storing.
\item We removed capitalization where it is a grammatical error.
\item After every i.e. there is a comma.
\item We are using vector spaces instead of finitely generated modules. There is only once where we still use finitely generated modules. This is because we want to store the information of the module (inside {\it Macaulay2}) that is used to construct the code.
\item We are using {\tt typewriter} font in the documentation where appropriated.
\item We updated the description of {\tt messages}.
\item We replaced the sentence ``For the best of our knowledge" for ``To the best of our knowledge".
\item We replaced the sentence ``the algorithm is well-implemented ``for "the algorithm is implemented well".
\item We replaced the sentence ``an error on the implementation" for ``an error in the implementation".
\item We replaced the sentence ``If generator or parity check matrix is not full rank" for ``If the generator matrix or the parity check matrix is not of full rank".
\item We replaced the sentence ``compute the equivalent reduce matrix" for ``Returns the reduced matrix".
\item We have put periods at ends of sentence of the code, we have also removed periods from the documentation headlines.
\item Fixed the punctuations and spacing.
\item We changed {\tt reduceMatrix} by {\tt reducedMatrix}.
\item We fixed ``repetition'' code.
\item We changed the names {\tt RMCode}, {\tt RSCode}, {\tt vasFunction}, {\tt hYpFunction} and {\tt gMdFunction} by {\tt reedMullerCode}, {\tt reedSolomonCode}, {\tt vasconcelosDegree }, {\tt hYp} and {\tt genMinDisIdeal}, respectively.
\end{itemize}

We hope that you find this revision acceptable.

\end{document}

### 2021-05-14 tex/answers.tex
Changed wording of file answers.tex and added info about the function minimumWeight

@@
 \end{center}
-We would like to thank you for your careful reading the manuscript and providing insightful comments.
-They have served to improve the submission.
-The first main change is the documentation in the package.
-We extended/improved/homogenized/latex-it all the documentation.
-The second main change is the minimum distance function, we found a flaw that \textcolor{red}{Nathan, write something please}.
-Here is a list of answers to the reviewer's comments that we performed during the preparation of this revision.
+We would like to thank you for your careful reading of our manuscript and for your insightful comments which have served to improve our submission.
+This document contains an overview of the changes that have been made in response to previous comments.
+\\
+
+In this new version, a number of documentation bugs have been fixed.
+Specifically, various instances of typos, lack of LaTex usage, poor wording and missing information have been addressed.
+We have also addressed various issues with code quality by refactoring various parts of the code.
+Additionally, this new iteration of the {\ttfamily CodingTheory} package addresses a number of bugs that were not mentioned by either of the reviewers.
+\\
+
+Here is a more specific list of responses to your comments:
 \begin{center}
@@
 \item[\rm 1.] Example 4.1 now contains the output of the command {\tt C.VanishingIdeal}.
-\item[\rm 2.] \textcolor{red}{Nathan, please write something about the minimum weight}
+\item[\rm 2.] As was suggested, an option to manually specify the strategy used by the function {\ttfamily minimumWeight} has been implemented.
+We have also made another change to this function which was not suggested in order to address apparent problems with performance.
+Namely, a new strategy called {\ttfamily OneInfoSet} which is strictly faster than the {\ttfamily BruteForce} strategy that it used previously has been added.
 \item[\rm 3.] We were not aware of this convention, thank you for pointing it out. We changed the name of the functions.

### 2021-05-15 tex/answers.tex
Changed optional argument 'Strategy' of minimumWeight to 'Strat' as a work-around for a bug causing documentation related warnings on installation.
added line about the GPLv3 license

@@
 We would like to thank you for your careful reading of our manuscript and for your insightful comments which have served to improve our submission.
-This document contains an overview of the changes that have been made in response to previous comments.
+This document contains an overview of the changes that we have made.
 \\
@@
 \item The package now works with {\it Macaulay2} 1.17.
-\item About public domain \textcolor{red}{Gwin, help please}
+\item Added a GPLv3 license header. 
 \item We have updated {\tt PackageImports} and {\tt PackageExports}.
@@
 \begin{itemize}
-\item There are no more warnings about missing documentation. \textcolor{red}{Nathan, there is a missing documentation now about the weight function}
+\item There are no more warnings about missing documentation.
 \item We corrected all the typos that you found. And more! We really want to thank you for all the comments, they were the basis to redo the documentation.

### 2021-05-22 tex/paper.tex
progess on cover letter/ paper

@@
 \end{example}
-
 \section{Basic parameters of linear codes}~\label{pro}
-
-The {\it dimension\/}, {\it length\/}, and {\it rate} or {\it information rate}  of a linear code are basic parameters.
-Given a linear code $C\subseteq \mathbb{F}^n_q$, these parameters are defined as $k=\dim_{\mathbb{F}_q} C$, $n$, and $k/n$ respectively.
-Another parameter of a linear code $C\subseteq \mathbb{F}^n_q$ is the {\it minimum Hamming weight}, which is given by
+The dimension $k$ and length $n$ are basic parameters of a linear code $C\subseteq \mathbb{F}^n_q$.
+The {\it information rate} of a linear code is defined as $k/n$.
+Another parameter of a linear code $C\subseteq \mathbb{F}^n_q$ is the {\it minimum distance}, which is defined as follows:
 \[
-w_H(C)=\min\{\|c\|
+w_H(C):=\min\{\|c\|
 \colon  c \in C, c \neq 0\},
 \]
-where $\| c \|$ is the number of non-zero entries of $c\in \mathbb{F}^n_q$.
-This parameter is important in determining the error-correcting capability of $C$; the higher the minimum Hamming weight, the more errors the code can detect and correct (see \cite{huf-pless}).
-This function has been implemented \textcolor{red}{Nathan, write here something about the implementation of the weight function}.
+where $\| c \|$ denotes the Hamming weight of $c\in \mathbb{F}^n_q$.
+This parameter is important in determining the error-correcting capability of $C$; the higher the minimum distance, the more errors the code can detect and correct (see \cite{huf-pless}).
+While the dimension and the length of linear code are computationally easy to determine, it is known that computing the minimum distance of an arbitrary linear code is an NP-hard problem \cite{Vardy}.
+For this task, the function {\ttfamily minimumWeight} is provided.
 
-In {\it Macaulay2}, the space $\mathbb{F}_q^n$ has been called the {\it ambient module} or {\it ambient space} of $C$, while the field $\mathbb{F}_q$, the {\it alphabet} or {\it field} of $C$.
+In {\it Macaulay2}, the space $\mathbb{F}_q^n$ has been called the {\it ambient module} or {\it ambient space} of $C$.
+The field $\mathbb{F}_q$ is called the {\it alphabet} or {\it field} of $C$.
 As often, the elements of $C$ are referred to as {\it codewords}.
-The input in all the following commands implemented in the {\tt CodingTheory} package is always a linear code  {\tt C}.
+The input in all the following commands implemented in the {\tt CodingTheory} package is always a linear code  {\tt C}:
 
@@
 
+\bibitem{Vardy} A.
+Vardy, "The intractability of computing the minimum distance of a code," in IEEE Transactions on Information Theory, vol.
+43, no.
+6, pp.
+1757-1766, Nov.
+1997, doi: 10.1109/18.641542.
+
 \bibitem{carvalho4} C. Carvalho, V. G. Neumann and H. H. L\'opez,

### 2021-05-23 tex/paper.tex

@@
 We implement functions to generate important families of linear codes such as Hamming codes, cyclic codes, Reed--Solomon codes, Reed--Muller codes, Cartesian codes, monomial--Cartesian codes, and toric codes.
-In addition, we implement functions for the syndrome decoding algorithm and locally recoverable code construction, which are important tools in applications of linear codes.
-The package \textit{CodingTheory} is available at \url{https://github.com/Macaulay2/Workshop-2020-Cleveland/tree/CodingTheory/CodingTheory}
+In addition, we implement functions for the syndrome decoding algorithm and locally recoverable code construction, which are important tools in applications of linear codes.% The package \textit{CodingTheory} is available at \url{https://github.com/Macaulay2/Workshop-2020-Cleveland/tree/CodingTheory/CodingTheory}
 \end{abstract}
@@
   \item {\tt E}, a list of elements of {\tt F}
-  \item {\tt L}, a list of vectors in $F^{\mathtt{m}}$
+  \item {\tt L}, a list of vectors in ${\tt F}^{\mathtt{m}}$
  \end{itemize}
@@
 
-\bibitem{Vardy} A.
-Vardy, "The intractability of computing the minimum distance of a code," in IEEE Transactions on Information Theory, vol.
-43, no.
-6, pp.
-1757-1766, Nov.
-1997, doi: 10.1109/18.641542.
-
 \bibitem{carvalho4} C. Carvalho, V. G. Neumann and H. H. L\'opez,
 Projective Nested Cartesian Codes,
-Bull Braz Math Soc, New Series (2016).
+Bulletin of the Brazilian Mathematical Society, New Series, {\bf 48} (2017), no. 2, 283--302.
 
-\bibitem{CSTVV} S. M. Cooper, A. Seceleanu, S. O. Toh\v{a}neanu,
-M. Vaz Pinto and R. H. Villarreal, 
-Generalized minimum distance functions and algebraic invariants of
-Geramita ideals, Adv. in Appl. Math. {\bf 112} (2020), 101940.
+\bibitem{CSTVV} S. M. Cooper, A. Seceleanu, S. O. Toh\v{a}neanu, M. Vaz Pinto and R. H. Villarreal, 
+Generalized minimum distance functions and algebraic invariants of Geramita ideals,
+Advances in Applied Mathematics, {\bf 112} (2020), 101940.
 
@@
 \bibitem{GLS} L. Gold, J. Little, H. Schenck,
-{Cayley-Bacharach and evaluation codes on complete intersections},
- J. Pure Appl. Algebra  {\bf 196} (1) (2005) 91--99.
+Cayley-Bacharach and evaluation codes on complete intersections,
+Journal of Pure and Applied Algebra, {\bf 196} (2005), no. 1, 91--99.
 
@@
 
-\bibitem{Ha1} J. Hansen, { Toric Surfaces and Error--correcting
-    Codes} in Coding Theory, Cryptography, and Related Areas, Springer
- (2000), pp. 132-142.
+\bibitem{Ha1} J. Hansen,
+Toric Surfaces and Error--correcting Codes,
+in Coding Theory, Cryptography, and Related Areas, (2000), pp. 132-142. Springer, Berlin, Heidelberg.
 
@@
 
-\bibitem{LSc} J. Little, H. Schenck,
-{\em Toric surface codes and Minkowski sums},
-SIAM J. Discrete Math. {\bf  20}  (2006),  no. 4, 999--1014 (electronic).
+\bibitem{LSc} J. Little and H. Schenck,
+Toric surface codes and Minkowski sums,
+SIAM Journal on Discrete Mathematics, {\bf  20}  (2006),  no. 4, 999--1014.
 
@@
 Affine Cartesian codes,
-Des.\ Codes Cryptogr. \textbf{71}(1) (2014) 5--19.
+{Designs, Codes and Cryptography}, {\bf 71} (2014), no. 1, 5--19.
 
@@
 
+\bibitem{MPV2} J. Mart\'inez-Bernal, Y. Pitones and R. H. Villarreal,
+Minimum distance functions of complete intersections,
+Journal of Algebra and Its Applications, {\bf 17} (2018), no. 11, 1850204 (22 pages).
+
 \bibitem{MPV} J. Mart\'inez-Bernal, Y. Pitones and R. H. Villarreal,
 Minimum distance functions of graded ideals and Reed-Muller-type codes,
-J. Pure Appl. Algebra 221 (2017), 251--275.
-
-\bibitem{MPV2} J. Mart\'inez-Bernal, Y. Pitones and R. H. Villarreal,
-Minimum distance functions of complete intersections,
-J. Algebra Appl. 17 (2018), no. 11, 1850204 (22 pages).
+Journal of Pure and Applied Algebra. {\bf 221} (2017), no. 2, 251--275.
 
@@
 Algebraic methods for parameterized codes and invariants of vanishing ideals over finite fields,
-Finite Fields Appl. {\bf 17} (2011), no. 1, 81--104.
+Finite Fields and Their Applications, {\bf 17} (2011), no. 1, 81--104.
 
 \bibitem{Ru}D. Ruano, 
-{On the parameters of $r$-dimensional toric codes},
-Finite Fields and Their Applications {\bf 13} (2007), pp. 962--976.
+On the parameters of $r$-dimensional toric codes,
+Finite Fields and Their Applications, {\bf 13} (2007), no. 4, pp. 962--976.
 
@@
 
-\bibitem{SoSo} I.~Soprunov, J.~Soprunova, {Toric surface codes and Minkowski length of polygons}, SIAM J. Discrete Math. {\bf 23}, Issue 1, (2009) 384-400
+\bibitem{SoSo} I.~Soprunov, J.~Soprunova,
+Toric surface codes and Minkowski length of polygons,
+SIAM Journal on Discrete Mathematics, {\bf 23} (2009), no. 1, 384--400.
 
-\bibitem{So1} I. Soprunov,  {Toric complete intersection codes.} Journal of Symbolic Computation. {\bf 50}, (2013) 374--385.
+\bibitem{So1} I. Soprunov,
+Toric complete intersection codes,
+Journal of Symbolic Computation, {\bf 50} (2013), 374--385.
 
@@
 A Family of Optimal Locally Recoverable Codes,
-IEEE Transactions on Information Theory {\bf 60} (2014), no. 8, 4661--4676.
+IEEE Transactions on Information Theory, {\bf 60} (2014), no. 8, 4661--4676.
 
 \bibitem{renteria-tapia-ca2} C. Renter\'\i a and H. Tapia-Recillas, 
-Reed-Muller codes: an ideal theory approach, Comm. Algebra {\bf 25}
-(1997), no. 2, 401--413.
+Reed-Muller codes: an ideal theory approach,
+Communications in Algebra, {\bf 25} (1997), no. 2, 401--413.
 
 \bibitem{van-lint} J. H. Van Lint,
-{ Introduction to coding theory},
+Introduction to coding theory,
 Third edition, Graduate Texts in Mathematics {\bf 86}, Springer-Verlag, Berlin, 1999.
+
+\bibitem{Vardy} A. Vardy,
+The intractability of computing the minimum distance of a code,
+IEEE Transactions on Information Theory, {\bf 43} (1997), no. 6, 1757--1766.
 
@@
 Monographs and Research Notes in Mathematics, Chapman and Hall/CRC,
-Boca Raton, FL, 2015.
+Boca Raton, FL (2015).
 \end{thebibliography}
